Flowchart of Current Algorithm

### 2008-10-24 docs/flowchart/flowchart.tex
more macro consolidation

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\ubold$     & total velocity ($\ubold = \Ut + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\Ut$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \Ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
\end{equation}

@@

\item Update the species.  Here we do a difference approximation
  to 
  \begin{equation}
  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\ubold \rho X_k) =
         \rho {\omegadot}_k \enskip , \label{eq:species}
  \end{equation}
  neglecting the reaction terms.  The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)$. 
  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  Taylor expansion in space and time, as described in paper II,
  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base
  state density, $\rho_0$, to convert this to an edge state for $\rho$,

@@

\item Update the enthalpy.  The full enthalpy equation is 
\MarginPar{we should say something about {\tt use\_tfromp}, and {\tt do\_eos\_h\_above\_cutoff}}
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
+ \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
\enskip , \label{eq:rhohupd} 
  \end{equation}
Here we consider only the advection, neglecting the reaction terms and thermal diffusion terms,
and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
+ \frac{Dp_0}{Dt} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h).$  Here we predict $(\rho
    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order Taylor expansion in space
    and time, as described in paper II, Appendix A, using $\V =
    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  We use the base state enthalpy,
    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
    \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}
\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
\enskip .
\end{equation}

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
    the conservative update of $(\rho X_k)$.  Here we predict $\rho'$ 
    and $X_k$ to the edges to obtain $\rho^{'(1),\nph}$ and 
    $X_k^{(1),\nph}$ using a second-order Taylor expansion in
    space and time, as described in paper II, Appendix A, using $\V =
    \uadvtwo+w_0^{\nph} \er$.  We use the base
    state density, $\rho_0$, to convert this to an edge state for $\rho$,

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
    using a second-order Taylor expansion in space and
    time, as described in paper II, Appendix A, using $\V =
    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
    terms in the enthalpy prediction since we accounted for them
    already in {\bf React State}. We use the base state enthalpy,
    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
\begin{equation}
(\rho h)^{(1),\nph} = 
(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
\end{equation}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\Ut^n$ to $\Ut^{n+1,\star}$ by discretizing
\begin{equation}
\frac{\partial\Ut}{\partial t} = - \Ut \cdotb \nablab \Ut - w_0 \frac{\partial \Ut}{\partial r}
                                 - \left(\Ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 ,
\end{equation}
(paper III; eq.~[37]) as 
\begin{eqnarray}
\Ut^{n+1,\star} &=& \Ut^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ut ]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\Ut^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
\enskip,
\end{equation}
by defining ${\bf V} = \Ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\Ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from
\begin{equation}
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\end{equation}

@@

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\Ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

@@

Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
Use the equation of state to determine $(\rho h)^{\initp}$.  Set
$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
proceeds as follows.

@@

  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$

@@

  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
        This yields $\ubold^{0,\nu+1}.$

@@

  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\Ut^{1,\star} - \Ut^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}

>>>>

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
\end{equation}

@@

\item Update the species.  Here we do a difference approximation
  to 
  \begin{equation}
  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) =
         \rho {\omegadot}_k \enskip , \label{eq:species}
  \end{equation}
  neglecting the reaction terms.  The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)$. 
  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  Taylor expansion in space and time, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
  state density, $\rho_0$, to convert this to an edge state for $\rho$,

@@

\item Update the enthalpy.  The full enthalpy equation is 
\MarginPar{we should say something about {\tt use\_tfromp}, and {\tt do\_eos\_h\_above\_cutoff}}
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
\enskip , \label{eq:rhohupd} 
  \end{equation}
Here we consider only the advection, neglecting the reaction terms and thermal diffusion terms,
and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \frac{Dp_0}{Dt} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h).$  Here we predict $(\rho
    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order Taylor expansion in space
    and time, as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  We use the base state enthalpy,
    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
    \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}
\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
\enskip .
\end{equation}

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
    the conservative update of $(\rho X_k)$.  Here we predict $\rho'$ 
    and $X_k$ to the edges to obtain $\rho^{'(1),\nph}$ and 
    $X_k^{(1),\nph}$ using a second-order Taylor expansion in
    space and time, as described in paper II, Appendix A, using $\Vb =
    \uadvtwo+w_0^{\nph} \er$.  We use the base
    state density, $\rho_0$, to convert this to an edge state for $\rho$,

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
    using a second-order Taylor expansion in space and
    time, as described in paper II, Appendix A, using $\Vb =
    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
    terms in the enthalpy prediction since we accounted for them
    already in {\bf React State}. We use the base state enthalpy,
    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
\begin{equation}
(\rho h)^{(1),\nph} = 
(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
\end{equation}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing
\begin{equation}
\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 ,
\end{equation}
(paper III; eq.~[37]) as 
\begin{eqnarray}
\Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ubt \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\Ubt^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\Ubt^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from
\begin{equation}
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\end{equation}

@@

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\Ubt^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

@@

Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
initial base state, and an initial guess for the velocity, $\Ub^{\initp}.$
Use the equation of state to determine $(\rho h)^{\initp}$.  Set
$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^{\initp}$ and 
$S = \rho\Hext$, giving $\Ub^{0,1}$.  The next part of the initialization process 
proceeds as follows.

@@

  \item Estimate $\Delta t^\nu$ using $\Ub^{0,\nu}$ and $w_0^\nu.$

@@

  \item Project $\Ub^{0,\nu}$ using $\beta_0^{\initp}$ and 
        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
        This yields $\Ub^{0,\nu+1}.$

@@

  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\Ub^0 = \Ub^{0,N_{\rm iters}^S+1}.$

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\Ubt^{1,\star} - \Ubt^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}
